### The purpose of these analyses is to evaluate activity measured by the number of steps taken per day over 5-minute intervals
###One analysis will evaluate if there are differences in the activity between week days and weekends

## Draw a histogram of the total number of steps per day
```{r setwde}
library(knitr)
setwd("C:/MyDocuments/Coursera/Data Science/Course 5 Reproducible Research/Week 2/Assign")

fulldata <- read.csv("activity.csv")

#Ignore all the NA values

nona <- na.omit(fulldata) 
nona$date <- factor(nona$date)

#Calculate the total number of steps per day
totsteps <- tapply(nona$steps,nona$date,sum,na.rm=TRUE)
totsteps <- data.frame(date=names(totsteps),steps=totsteps)
#Draw a histogram of the total number of steps per day
#plot(totsteps$date,totsteps$steps, main="Total Steps Per Day", type="h",xlab="Date",ylab="Steps")
hist(totsteps$steps, col = "grey",breaks = 16,ylim = c(0,10),xlim = c(0,25000)
                                        ,xlab = "Steps", main = "Total Steps Per Day")

#Calculate the mean number of steps taken per day
stepsmean <- format(mean(totsteps$steps,na.rm=TRUE),digits=9)
#Calculate the median number of steps taken per day
stepsmedian <- format(median(totsteps$steps,na.rm=TRUE),digits=9)
```

## Mean and Median
The mean is `r stepsmean`, and the median is `r stepsmedian`.

## Steps per Interval
```{r interval_mean}
#Calculate the mean steps per interval
avgsteps <- tapply(nona$steps,nona$interval,mean,na.rm=TRUE)
avgsteps <- data.frame(intervals=names(avgsteps),values=avgsteps)
avgsteps$intervals <- strptime(avgsteps$intervals,"%H%M")

#Draw the average steps per interval on a time series plot
plot(avgsteps$intervals,avgsteps$values, main="Average Steps Per Interval", type="l",xlab="Interval",ylab="Steps")

#Determine which interval contains the highest number of steps
moststeps <- row.names(avgsteps)[which(avgsteps$values==max(avgsteps$values))]
```

The most number of step occurs in interval `r moststeps`

## NA Values
```{r navals}
#Calculate the total number of missing values in the original data
numna <- nrow(fulldata[fulldata$steps=="NA",])
```

There are `r numna` rows with NA values

```{r replacena}
#Replace NA values with the average of the interval
for (i in 1:(nrow(fulldata))) {
    if (is.na(fulldata$steps[i])) 
    {
        fulldata$steps[i] <- avgsteps$values[match(fulldata$interval[i], avgsteps$interval)]
        ##print(paste(i,fillna$steps[i],sep=" "))
    }
}
```
```{r calcsteps}
#Calculate the total number of steps per day
totsteps2 <- tapply(fulldata$steps,fulldata$date,sum,na.rm=TRUE)
totsteps2 <- data.frame(date=names(totsteps2),steps=totsteps2)

#Draw a histogram of the total number of steps per day
plot(totsteps2$date,totsteps2$steps, main="Total Steps Per Day NA`s replaced", type="h",xlab="Interval",ylab="Steps")

#Calculate the mean number of steps taken per day
stepsmean2 <- format(mean(totsteps2$steps,na.rm=TRUE),digits=9)
#Calculate the median number of steps taken per day
stepsmedian2 <- format(median(totsteps2$steps,na.rm=TRUE),digits=9)
```

## Mean and Median (NA values removed)
With no NA values, the mean is `r stepsmean2`, and the median is `r stepsmedian2`.

Both the mean and median have increased when the NA values were replaced by the average per interval

## Weekday vs Weekend Analysis
```{r week}
#Create a new factor variable to indicate if the date is a WEEKDAY or WEEKEND
for (i in 1:(nrow(fulldata))) {
if (weekdays(as.Date(fulldata$date[i],abbreviate=FALSE)) == 'Saturday' | weekdays(as.Date(fulldata$date[i],abbreviate=FALSE)) == 'Sunday') 
{
fulldata$week[i] <- 'WEEKEND'
#print(paste(i,fulldata$week[i],sep=" "))
}
else 
{
fulldata$week[i] <- 'WEEKDAY'
}
}

#Split the dataset into 2 seperate datasets to plot WEEKDAY and WEEKEND on seperate plots
weekdata <- subset(fulldata,fulldata$week == "WEEKDAY")
weekenddata <- subset(fulldata,fulldata$week == "WEEKEND")

totweekdata <- tapply(weekdata$steps,weekdata$interval,sum,na.rm=TRUE)
totweekdata <- data.frame(interval=names(totweekdata),steps=totweekdata)

totweekenddata <- tapply(weekenddata$steps,weekenddata$interval,sum,na.rm=TRUE)
totweekenddata <- data.frame(interval=names(totweekenddata),steps=totweekenddata)

#Calculate the y-axis range
rng <- range(totweekdata$steps,totweekenddata$steps)

par(mfrow=c(2,1), mar=c(4,4,2,1)) # set to disply plot in 2 rows with 1 colmn and adjust margins
plot(totweekdata$interval, totweekdata$steps,type="l", main="Weekdays", ylim=rng,xlab="Interval",ylab="Steps")
plot(totweekenddata$interval, totweekenddata$steps,type="l", main="Weekends", ylim=rng,xlab="Interval",ylab="Steps")
```
